---
title: "Lab 01 Exploring Data by- Sachin Mohan Sujir"
output:
  html_document:
    df_print: paged
---

# (a) 

Reading csv file

```{r a}
rm(list=ls())
College=read.csv("College.csv",na.strings="?",header = T)
```

# (b) 

Viewing Data and Selecting first column

```{r b}
fix(College)
rownames(College) = College[,1]
fix(College)
```


Eliminating the first column

```{r b1}
College = College[,-1]
fix(College)
```

# (c)

(i)Summary Function

```{r c}
summary(College)
```

(ii) Using pair to display scatter plot for first 10 columns

```{r c1}
pairs(College[,1:10])
```

(iii) plot() function to produce side-by-side boxplots of Outstate versus
Private

```{r c2}
College$Private = as.factor(College$Private)
plot(College$Private,College$Outstate,xlab="Private",ylab= "Outstate",main="Boxplot: Outstate vs Private")
```

(iv) Create a new qualitative variable, called Elite, by binning the Top10perc variable, divide universities into two groups based on whether or
not the proportion of students coming from the top 10% of their high school
classes exceeds 50%.

```{r c3}
Elite = rep("No",nrow(College))
Elite[College$Top10perc > 50] = "Yes"
Elite = as.factor(Elite)
College = data.frame(College,Elite)
summary(College)

```

Elite Universities: 78

Box Plot Outsource vs Elite

```{r c4}
College$Elite = as.factor(College$Elite)
plot(College$Elite,College$Outstate,xlab="Elite",ylab= "Outstate",main="Boxplot: Outstate vs Elite")
```

(v)Histogram for different variables

```{r c5}
par(mfrow=c(2,2))
hist(College$Enroll,col="Red")
hist(College$Accept,col="Green")
hist(College$Apps,col="Yellow")
hist(College$PhD,col="Blue")
```

# (d)
(i) Which of the predictors are quantitative, and which are qualitative?

```{r}
sapply(College,class)
dim(College)
```

From the above snippet we can see that Private and Elite are the only Qualitative Variables(Factor). Rest of the Variables are Quantitative(numberic). So, 2 Qualitative and 17 Quantitative Variables.

(ii) Range, Mean, Standard Deviation of Enroll

```{r}
rangeEnroll=range(College$Enroll)
meanEnroll=mean(College$Enroll)
sdEnroll=sd(College$Enroll)
"Range of Enroll: " 
rangeEnroll

"Mean of Enroll :"
meanEnroll

"Standard Deviation of Enroll: "
sdEnroll
```

(iii) Removing 100th through 200th observations and finding range, sd and mean of Enroll

```{r}
newenroll = College$Enroll[c(-100,-200)]
rangeEnroll_1=range(newenroll)
meanEnroll_1=mean(newenroll)
sdEnroll_1=sd(newenroll)
"Range of Enroll: " 
rangeEnroll_1

"Mean of Enroll :"
meanEnroll_1

"Standard Deviation of Enroll: "
sdEnroll_1

```

Comparing the range, mean and sd of new enroll with original enroll in college.csv, we see that they remain unchanged.



